Pool Model - A unified Mathematical\\ Modelling Framework for Food Science

### 2025-09-05 paper/pool_model.tex
write discussion
- also remove bullet points

How well does the pool model do to encapsulate all the different growth models
\begin{itemize}
    \item All 6 models can be represented by pool model
    \item Baranyi-Roberts results can be reproduced, but not identical; requires some work
    \item what are some general limitations within this framework? how can they be circumvented?
    \item why is the pool model "better" than the other \ac{ode} models? (why did we do this?);
        generalization of many models; better explainability;
    \item link to supplement; we can generalize this and prove theorems for the abstract,
        generalized model; this allows us to discuss many cases at once
\end{itemize}

How justified was the construction of the \ac{abm}?
\begin{itemize}
    \item Discuss simplified physical interaction model
    \item Heterogeneity plays a role; as we have shown
    \item why did we not simply use a PDE?; because we have to think in terms of cellular processes
        and this ensures that our model is actually realistic
\end{itemize}

How can we cover more effects of bacteria?
\begin{itemize}
    \item More states for metabolic processes
    \item More states for more species
    \item Other cooperiation/inhibition effects
    \item external influences
\end{itemize}

What are limitations of our approach?
\begin{itemize}
    \item spatial stuff; heterogeneity (because \ac{ode})
    \item larger models require more parameters, harder to estimate
    \item only simple models have analytical solutions
\end{itemize}

>>>>

% How well does the pool model do to encapsulate all the different growth models
% \begin{itemize}
%     \item All 6 models can be represented by pool model
%     \item Baranyi-Roberts results can be reproduced, but not identical; requires some work
%     \item what are some general limitations within this framework? how can they be circumvented?
%     \item why is the pool model "better" than the other \ac{ode} models? (why did we do this?);
%         generalization of many models; better explainability;
%     \item link to supplement; we can generalize this and prove theorems for the abstract,
%         generalized model; this allows us to discuss many cases at once
% \end{itemize}
%
% How can we cover more effects of bacteria?
% \begin{itemize}
%     \item More states for metabolic processes
%     \item More states for more species
%     \item Other cooperiation/inhibition effects
%     \item external influences
% \end{itemize}
%
% How justified was the construction of the \ac{abm}?
% \begin{itemize}
%     \item Discuss simplified physical interaction model
%     \item Heterogeneity plays a role; as we have shown
%     \item why did we not simply use a PDE?; because we have to think in terms of cellular processes
%         and this ensures that our model is actually realistic
% \end{itemize}
%
% What are limitations of our approach?
% \begin{itemize}
%     \item spatial stuff; heterogeneity (because \ac{ode})
%     \item larger models require more parameters, harder to estimate
%     \item only simple models have analytical solutions
% \end{itemize}

The pool model presented here was used to describe the dynamics of 6 systems containing established models of bacterial growth.
Comparing our approach with the Baranyi-Roberts model which is widely used in food science, it is possible to obtain results which provide growth curves of very similar shape.
However, their analytical form is not completely identical.
The inclusion of additional metabolic processes or larger number of cell-species can be achieved by extending the list of possible states and defining transitions between them.
In addition we have shown how to treat interspecies interactions in the form of cooperation and competition.
This framework provides a starting point for a mathematically more abstract formulation which is thus able to cover a much wider class of systems.
By constructing an \ac{abm}, we have exemplified how close the devised model is to the individual biological processes that occur within the bacteria.
We have investigated the effect of spatial effects in the form of varying diffusion parameters and different initial spatial distributions of bacteria.
Our findings reinforce that the \ac{ode} model is a good simplification in the highly-diffusive regime within a mostly homogenous spatial bacterial distribution.
However, deviations to these assumptions can lead to non-negligible effects, influencing the growth of the colony.
